---
title: "Portfolio Optimization"
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), paste0("Reports/PortfOpt_",lubridate::today(),"_week-",lubridate::week(lubridate::today()),".html"))) })
params:
  date: !r Sys.Date()
---


```{r "libraries", echo=FALSE, results='hide', warning=FALSE,message=FALSE}

report_start_time <- Sys.time()
setwd("D:/Project/DataScience/Portfolio-optimization")
suppressPackageStartupMessages(library(tidyquant)) # To download the data
suppressPackageStartupMessages(library(plotly)) # To create interactive charts
suppressPackageStartupMessages(library(timetk)) # To manipulate the data series
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(PortfolioAnalytics))
suppressPackageStartupMessages(library(ggrepel))
options(scipen=999)
```

```{r "key variables", echo = FALSE}
num_port <- 200000
num_fit <- 100000

wts_have <- c("AMD" = 0.3088,
              "TAL" = 0.1169,
              "FRHC" = 0.4232,
              "PLZL.ME" = 0.0891,
              "KR" = 0.0219,
              "LSCC" = 0.0398
              )
```

```{r "proceeding", echo=FALSE, results='hide', warning=FALSE,message=FALSE}

tick <- c("AMD", "NVTA",
          "FRHC", "TAL", "NOK", "FCN", "FRPT", "APPS", "MRTX", "ZYXI", "PLZL.ME",
          "PWR", "GSHD", "WPM", "MOEX.ME")

tick <- unique(c(tick, names(wts_have)))


price_from <- "2017-12-27"
price_to <- "2021-12-27"

price_data <- tq_get(tick,
                     from = price_from,
                     to = price_to,
                     get = 'stock.prices')

# convert russian assets to dollars

usdrub <- tq_get("RUB=X", from = price_from,
                 to = price_to,
                 get = 'stock.prices')
usdrub <- select(usdrub, date, adjusted)

price_data <- price_data %>%
    left_join(usdrub, by = "date", suffix = c(".stock", ".dollar")) %>%
    mutate(adjusted = if_else(grepl(".ME", price_data$symbol),
                              adjusted.stock / adjusted.dollar, 
                              adjusted.stock)) %>%
    select(-c(adjusted.dollar, adjusted.stock))

# end converting

log_ret_tidy <- price_data %>%
    group_by(symbol) %>%
    tq_transmute(select = adjusted,
                 mutate_fun = periodReturn,
                 period = 'daily',
                 col_rename = 'ret',
                 type = 'log')

log_ret_xts <- log_ret_tidy %>%
    pivot_wider(names_from = symbol, values_from = ret) %>%
    tk_xts()

log_ret_xts_compl <- na.omit(log_ret_xts) 



cov_mat <- cov(log_ret_xts_compl) * 252

#round(cor(log_ret_xts_compl),2)
# print(round(cov_mat,4))

# Creating a matrix to store the weights

all_wts <- matrix(nrow = num_port,
                  ncol = length(tick))

# Creating an empty vector to store
# Portfolio returns

port_returns <- vector('numeric', length = num_port)

# Creating an empty vector to store
# Portfolio Standard deviation

port_risk <- vector('numeric', length = num_port)

# Creating an empty vector to store
# Portfolio Sharpe Ratio

sharpe_ratio <- vector('numeric', length = num_port)

risk_free <- 0.05


```

```{r "portfolio fit", echo=FALSE, message=FALSE, warning=FALSE, results= FALSE}
# Simulation

sharpe_ratio_fit <- 0
cnt_tries <- 0
cnt_info <- 0
cnt_total <- 0
start_fit <- Sys.time()
while (cnt_tries <= num_fit) {
    cnt_tries <- cnt_tries + 1
    cnt_info <- cnt_info + 1
    wts <- runif(length(tick))
    wts <- round(wts/sum(wts),2)
    
    # Storing weight in the matrix
    #all_wts[i,] <- wts
    
    # Portfolio returns
    
    #port_fit_ret <- sum(wts * mean_ret)
    #port_fit_ret <- ((port_fit_ret + 1)^252) - 1
      
    port_fit_ret <- log_ret_xts_compl %*% wts
    port_fit_ret <- prod(port_fit_ret + 1)^(252/nrow(port_fit_ret)) - 1

    # Creating and storing portfolio risk
    port_fit_sd <- sqrt(t(wts) %*% (cov_mat  %*% wts))
    
    
    # Creating and storing Portfolio Sharpe Ratios
    # Assuming 0% Risk free rate
    
    sr <- (port_fit_ret-risk_free)/port_fit_sd
    if(round(sr,3) > sharpe_ratio_fit){
        cnt_tries <- 0
        cnt_info <- 0
        wts_fit <- wts
        port_fit_returns <- round(port_fit_ret,3)
        port_fit_risk <- round(port_fit_sd,3)
        sharpe_ratio_fit <- round(sr,3)
        time_past <- round(difftime(Sys.time(), start_fit, 
                                    units = "mins"), 2)
        message("Sharpe ratio: ",sharpe_ratio_fit,
            " | Time past: ", time_past, " min", "\r")
    }
    if(cnt_info == 1000000){
        cnt_info <- 0
        time_past <- round(difftime(Sys.time(), start_fit, 
                                    units = "mins"), 2)
        message("Tries: ", cnt_tries/1000000, " M", 
            " | Sharpe ratio: ", sharpe_ratio_fit, 
            " | Time past: ", time_past, " min", "\r")
        
    }
    cnt_total <- cnt_total + 1
}
finish_fit <- Sys.time()
spent <- difftime(finish_fit, start_fit, units = "mins")

# Storing the values in the table
fit_portfolio_values <- tibble(Return = port_fit_returns,
                           Risk = port_fit_risk,
                           SharpeRatio = sharpe_ratio_fit)


# Converting matrix to a tibble and changing column names
wts_fit <- as_tibble(rbind(wts_fit))
colnames(wts_fit) <- colnames(log_ret_xts)

# Combing all the values together
fit_portfolio_values <- tk_tbl(cbind(wts_fit, fit_portfolio_values), preserve_index = F)
```

Total fit trials : `r round(cnt_total/1000000)`M

```{r "simulation", echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# Simulation
start <- Sys.time()
for (i in seq_along(port_returns)) {
    
    wts <- runif(length(tick))
    wts <- round(wts/sum(wts),2)
    
    # Storing weight in the matrix
    all_wts[i,] <- wts
    
    # Portfolio returns
    
    #port_ret <- sum(wts * mean_ret)
    #port_ret <- ((port_ret + 1)^252) - 1
    
    port_ret <- log_ret_xts_compl %*% wts
    port_ret <- prod(port_ret + 1)^(252/nrow(port_ret)) - 1
    
    # Storing Portfolio Returns values
    port_returns[i] <- round(port_ret,3)
    
    
    # Creating and storing portfolio risk
    port_sd <- sqrt(t(wts) %*% (cov_mat  %*% wts))
    port_risk[i] <- round(port_sd,3)
    
    # Creating and storing Portfolio Sharpe Ratios
    # Assuming 0% Risk free rate
    
    sr <- (port_ret-risk_free)/port_sd
    sharpe_ratio[i] <- round(sr,3)
}
finish <- Sys.time()
spent <- difftime(finish, start, units = "mins")

# Storing the values in the table
portfolio_values <- tibble(Return = port_returns,
                           Risk = port_risk,
                           SharpeRatio = sharpe_ratio)


# Converting matrix to a tibble and changing column names
all_wts <- tk_tbl(all_wts)

colnames(all_wts) <- colnames(log_ret_xts)

# Combing all the values together
portfolio_values <- tk_tbl(cbind(all_wts, portfolio_values))


portfolio_values <- portfolio_values %>%
    distinct(Risk, Return, .keep_all = TRUE)
```

Time spent on simulation: `r round(spent)` minutes

Portfolios amount: `r as.character(num_port/1000000)`M

Date range from `r min(index(log_ret_xts_compl))` to `r max(index(log_ret_xts_compl))`

Risk free: `r risk_free`

```{r "single asset", echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# Simulation

wts_singles <- matrix(rep(0, length(tick)^2), 
                      nrow = length(tick), ncol = length(tick))
diag(wts_singles) <- 1

port_singles_returns <- vector('numeric', length = length(tick))

port_singles_risk <- vector('numeric', length = length(tick))

sharpe_ratio_singles <- vector('numeric', length = length(tick))

for (i in 1:length(tick)) {
    # Portfolio returns
    
    #port_ret <- sum(wts_singles[i,] * mean_ret)
    #port_ret <- ((port_ret + 1)^252) - 1
    
    port_ret <- log_ret_xts_compl %*% wts_singles[i,]
    port_ret <- abs(prod(port_ret + 1))^(252/nrow(port_ret)) - 1
    
    
    # Storing Portfolio Returns values
    port_singles_returns[i] <- round(port_ret,3)
    
    
    # Creating and storing portfolio risk
    port_singles_sd <- sqrt(t(wts_singles[i,]) %*% (cov_mat  %*% wts_singles[i,]))
    port_singles_risk[i] <- round(port_singles_sd,3)
    
    # Creating and storing Portfolio Sharpe Ratios
    # Assuming 0% Risk free rate
    
    sr <- (port_ret-risk_free)/port_singles_sd
    sharpe_ratio_singles[i] <- round(sr,3)
}

# Storing the values in the table
single_portfolio_values <- tibble(Return = port_singles_returns,
                           Risk = port_singles_risk,
                           SharpeRatio = sharpe_ratio_singles)


# Converting matrix to a tibble and changing column names
wts_singles <- tk_tbl(wts_singles)

colnames(wts_singles) <- colnames(log_ret_xts)

# Combing all the values together
single_portfolio_values <- tk_tbl(cbind(wts_singles, single_portfolio_values))

single_portfolio_values <- single_portfolio_values %>%
    pivot_longer(1:(ncol(single_portfolio_values)-3),
             names_to = "Asset") %>%
    filter(value == 1) %>%
    select(-value)

single_portfolio_values
    
```

## My current portfolio

```{r "my portfolio", echo = FALSE, warning = FALSE}
# Calculate the random weights

#wts_one <- c(0, 0.2751, 0.1958, 0.5290, rep(0,7))
#wts_one <- c(0.4583, 0.3162, 0.2255, rep(0,8))

wts_one <- rep(0, length(tick))
names(wts_one) <- tick
wts_one <- replace(wts_one, names(wts_have), wts_have)

# Calculate the portfolio returns
port_return_one <- log_ret_xts_compl %*% as.numeric(wts_one)
port_return_one <- prod(port_return_one + 1)^(252/nrow(port_return_one)) - 1



# Calculate the portfolio risk
port_risk_one <- sqrt(t(as.numeric(wts_one)) %*% (cov_mat %*% as.numeric(wts_one)))

# Calculate the Sharpe Ratio
sharpe_ratio_one <- (port_return_one-risk_free)/port_risk_one

portfolio_values_one <- tibble(Return = port_return_one,
                               Risk = port_risk_one,
                               SharpeRatio = sharpe_ratio_one)

# Converting matrix to a tibble and changing column names
wts_one <- as_tibble(rbind(wts_one))
colnames(wts_one) <- colnames(log_ret_xts)

# Combing all the values together
portfolio_values_one <- tk_tbl(cbind(wts_one, portfolio_values_one), preserve_index = F)

kbl(round(portfolio_values_one,2)) %>%
    kable_styling()
```
```{r "optimized_sharpie", echo=FALSE, message=FALSE, warning=FALSE, results= FALSE}
R <- log_ret_xts_compl

funds <- colnames(R)


# Construct initial portfolio
init.portf <- portfolio.spec(assets=funds)
init.portf <- add.constraint(portfolio=init.portf, type="full_investment")
init.portf <- add.constraint(portfolio=init.portf, type="long_only")
init.portf <- add.objective(portfolio=init.portf, type="return", name="mean")
init.portf <- add.constraint(portfolio=init.portf, type="position_limit", max_pos=7)
# init.portf <- add.objective(portfolio=init.portf, type="risk", name="StdDev")
init.portf <- add.objective(portfolio=init.portf, type='risk', name='ETL',
                            arguments=list(p=0.95))
maxSR.lo.ROI <- optimize.portfolio(R=R, portfolio=init.portf, 
                                   optimize_method="ROI", 
                                   maxSR=TRUE, trace=TRUE)
maxSRReturns <- Return.portfolio(R,weight = extractWeights(maxSR.lo.ROI), 
                                   rebalance_on = "days")
roi_wts <- as_tibble(t(extractWeights(maxSR.lo.ROI)))

opt_sr_res <- table.AnnualizedReturns(R = maxSRReturns, Rf = 0.05/252)
opt_sr_portfolio_values <- tibble(Return = opt_sr_res[1,],
                           Risk = opt_sr_res[2,],
                           SharpeRatio = opt_sr_res[3,])
opt_sr_portfolio_values <- cbind(roi_wts, opt_sr_portfolio_values)
kbl(round(opt_sr_portfolio_values,2)) %>%
    kable_styling()
```


## Minimum Variance Portfolio Weights

```{r "minimum variance portfolio", echo = FALSE, out.width='100%', warning=FALSE}
# The minimum variance portfolio
min_var <- portfolio_values[which.min(portfolio_values$Risk),]
p <- min_var %>%
    gather(names(min_var)[1]:names(min_var)[length(names(min_var))-3], 
           key = Asset,
           value = Weights) %>%
    mutate(Asset = as.factor(Asset)) %>%
    ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
    geom_bar(stat = 'identity') +
    geom_text(aes(label = round(Weights,2)), nudge_y = 0.02, size = 3.5,
                color = "darkgray") +
    theme_minimal() +
    labs(x = 'Assets', y = 'Weights', title = "Minimum Variance Portfolio Weights") +
    scale_y_continuous(labels = scales::percent) 

min_var <- min_var[,order(min_var[nrow(min_var),])]
min_var <- relocate(min_var, Return, Risk, SharpeRatio,.after = last_col())

kbl(round(min_var,2)) %>%
    kable_styling()


ggplotly(p)
```

## Tangency Portfolio Weights

```{r "tangency portfolio", echo = FALSE, out.width='100%', warning=FALSE}
max_sr <- portfolio_values[which.max(portfolio_values$SharpeRatio),]

p <- max_sr %>%
    gather(names(max_sr)[1]:names(max_sr)[length(names(max_sr))-3], 
           key = Asset,
           value = Weights) %>%
    mutate(Asset = as.factor(Asset)) %>%
    ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
    geom_bar(stat = 'identity') +
    geom_text(aes(label = round(Weights,2)), nudge_y = 0.02, size = 3.5,
                color = "darkgray") +
    theme_minimal() +
    labs(x = 'Assets', y = 'Weights', 
         title = "Tangency Portfolio Weights") +
    scale_y_continuous(labels = scales::percent)

max_sr <- max_sr[,order(max_sr[nrow(max_sr),])]
max_sr <- relocate(max_sr, Return, Risk, SharpeRatio,.after = last_col())

kbl(round(max_sr,2)) %>%
    kable_styling()
ggplotly(p)
```

## Portfolio Sharpie ratio fit

```{r "portfolio fit res", echo=FALSE, out.width='100%'}
p <- fit_portfolio_values %>%
    gather(names(fit_portfolio_values)[1]:names(fit_portfolio_values)[length(names(fit_portfolio_values))-3], 
           key = Asset,
           value = Weights) %>%
    mutate(Asset = as.factor(Asset)) %>%
    ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
    geom_bar(stat = 'identity') +
    geom_text(aes(label = round(Weights,2)), nudge_y = 0.02, size = 3.5,
              color = "darkgray") +
    theme_minimal() +
    labs(x = 'Assets', y = 'Weights', 
         title = "Fit Portfolio Weights") +
    scale_y_continuous(labels = scales::percent)

kbl(round(fit_portfolio_values,2)) %>%
    kable_styling()

ggplotly(p)

```

## Portfolio Optimized Shaprie ratio fit (ROI method)
```{r "optimized_sharpie res", echo = FALSE, out.width='100%', warning=FALSE}
kbl(round(opt_sr_portfolio_values,2)) %>%
    kable_styling()

p <- roi_wts %>%
    gather(names(roi_wts), 
           key = Asset,
           value = Weights) %>%
    filter(Weights > 0.001) %>%
    mutate(Asset = as.factor(Asset)) %>%
    ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
    geom_bar(stat = 'identity') +
    geom_text(aes(label = round(Weights,2)), nudge_y = 0.02, size = 3.5,
              color = "darkgray") +
    theme_minimal() +
    scale_y_continuous(labels = scales::percent) +
    labs(x = 'Assets', y = 'Weights', title = "Optimized Portfolio Weights",
         subtitle = paste("From:", min(index(R)), "to", max(index(R)),
                          "| SR:", round(SharpeRatio.annualized(maxSRReturns, 
                                                                Rf = 0.05/252)[1,1],2),
                          "| Return:", round(Return.annualized(maxSRReturns)[1,1],2),
                          "| Risk: ", round(StdDev.annualized(maxSRReturns)[1,1],2)))
# p <- p + theme(legend.position = "none")
ggplotly(p)

```


```{r "efficient frontier", echo = FALSE, out.width='100%'}
# Visualizing all the portfolios
risk_reward_port <- data.frame(risk = as.numeric(), return = as.numeric(), 
                               portfolio = character())
risk_reward_port[1,] <- c(min_var$Risk, min_var$Return, "Min Var")
risk_reward_port[2,] <- c(max_sr$Risk, max_sr$Return, "Max SR sim")
risk_reward_port[3,] <- c(fit_portfolio_values$Risk, fit_portfolio_values$Return,
                          "Max SR fit")
risk_reward_port[4,] <- c(portfolio_values_one$Risk, portfolio_values_one$Return,
                          "My portfolio")
risk_reward_port[5,] <- c(opt_sr_portfolio_values$Risk, opt_sr_portfolio_values$Return,
                          "SR ROI")
risk_reward_port$risk <- as.numeric(risk_reward_port$risk)
risk_reward_port$return <- as.numeric(risk_reward_port$return)
risk_reward_port$portfolio <- as.factor(risk_reward_port$portfolio)

p <- portfolio_values %>%
    ggplot(aes(x = Risk, y = Return)) +
    geom_point(alpha = 0.1, color = "gray30") +
    theme_classic() +
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(labels = scales::percent) +
    #coord_cartesian(xlim = c(min(portfolio_values$Risk) - 0.02,
    #                         max(portfolio_values$Risk) + 0.02),
    #                ylim = c(min(portfolio_values$Return - 0.02),
    #                         max(portfolio_values$Return) + 0.02)) +
    labs(x = 'Annualized Risk',
         y = 'Annualized Returns',
         title = "Portfolio Optimization & Efficient Frontier") +
    geom_point(aes(x = risk, y = return, color = portfolio), 
               data = risk_reward_port, size = 2) +
    geom_text_repel(aes(x = Risk, y = Return, label = Asset),
              data = single_portfolio_values, 
              color = "gray25", size = 3) +
    theme(legend.position="bottom")
p
```

Total time: `r round(difftime(Sys.time(), report_start_time, units = "secs"))` seconds
