---
title: "Portfolio Optimization"
output: html_document
---

```{r "libraries", echo=FALSE, results='hide', warning=FALSE,message=FALSE}
report_start_time <- Sys.time()
suppressPackageStartupMessages(library(tidyquant)) # To download the data
suppressPackageStartupMessages(library(plotly)) # To create interactive charts
suppressPackageStartupMessages(library(timetk)) # To manipulate the data series
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(kableExtra))
options(scipen=999)
```

```{r "key variables", echo = FALSE}
num_port <- 12000000
num_fit <- 70000000

wts_have <- c("AMD" = 0.3210,
              "AAPL" = 0.6696)
```


```{r "proceeding", echo=FALSE, results='hide', warning=FALSE,message=FALSE}
#tick <- c("AAPL", "AMD", "PLZL.ME", "SBER.ME", 'ALRS.ME', 'MTSS.ME'
#          , "YNDX.ME", "MOEX.ME", "NVDA", "TSLA", "BABA", "QCOM")

tick <- c("AAPL", "AMD", "PLZL.ME", "SBER.ME", "QCOM", "AXSM",
          "NVTA", "ACMR", "FATE", "POLY.ME")

price_from <- "2017-11-16"
price_to <- "2020-11-16"

price_data <- tq_get(tick,
                     from = price_from,
                     to = price_to,
                     get = 'stock.prices')

# convert russian assets to dollars

usdrub <- tq_get("RUB=X", from = price_from,
                 to = price_to,
                 get = 'stock.prices')
usdrub <- select(usdrub, date, adjusted)

price_data <- price_data %>%
    left_join(usdrub, by = "date", suffix = c(".stock", ".dollar")) %>%
    mutate(adjusted = if_else(grepl(".ME", price_data$symbol),
                              adjusted.stock / adjusted.dollar, 
                              adjusted.stock)) %>%
    select(-c(adjusted.dollar, adjusted.stock))

# end converting
price_data <- na.omit(price_data)

log_ret_tidy <- price_data %>%
    group_by(symbol) %>%
    tq_transmute(select = adjusted,
                 mutate_fun = periodReturn,
                 period = 'weekly',
                 col_rename = 'ret',
                 type = 'log')

log_ret_xts <- log_ret_tidy %>%
    pivot_wider(names_from = symbol, values_from = ret) %>%
    tk_xts()

log_ret_xts_compl <- na.omit(log_ret_xts) 

mean_ret <- colMeans(log_ret_xts_compl)
# print(round(mean_ret, 5))

cov_mat <- cov(log_ret_xts_compl) * 52

#round(cor(log_ret_xts_compl),2)
# print(round(cov_mat,4))

# Creating a matrix to store the weights

all_wts <- matrix(nrow = num_port,
                  ncol = length(tick))

# Creating an empty vector to store
# Portfolio returns

port_returns <- vector('numeric', length = num_port)

# Creating an empty vector to store
# Portfolio Standard deviation

port_risk <- vector('numeric', length = num_port)

# Creating an empty vector to store
# Portfolio Sharpe Ratio

sharpe_ratio <- vector('numeric', length = num_port)

risk_free <- 0.05


```



```{r "portfolio fit", echo=FALSE, message=FALSE, warning=FALSE, results= FALSE}
# Simulation

sharpe_ratio_fit <- 0
cnt_tries <- 0
cnt_info <- 0
cnt_total <- 0
start_fit <- Sys.time()
while (cnt_tries <= num_fit) {
    cnt_tries <- cnt_tries + 1
    cnt_info <- cnt_info + 1
    wts <- runif(length(tick))
    wts <- round(wts/sum(wts),2)
    
    # Storing weight in the matrix
    #all_wts[i,] <- wts
    
    # Portfolio returns
    
    port_fit_ret <- sum(wts * mean_ret)
    port_fit_ret <- ((port_fit_ret + 1)^52) - 1

    # Creating and storing portfolio risk
    port_fit_sd <- sqrt(t(wts) %*% (cov_mat  %*% wts))
    
    
    # Creating and storing Portfolio Sharpe Ratios
    # Assuming 0% Risk free rate
    
    sr <- (port_fit_ret-risk_free)/port_fit_sd
    if(round(sr,3) > sharpe_ratio_fit){
        cnt_tries <- 0
        cnt_info <- 0
        wts_fit <- wts
        port_fit_returns <- round(port_fit_ret,3)
        port_fit_risk <- round(port_fit_sd,3)
        sharpe_ratio_fit <- round(sr,3)
        time_past <- round(difftime(Sys.time(), start_fit, 
                                    units = "mins"), 2)
        message("Sharpe ratio: ",sharpe_ratio_fit,
            " | Time past: ", time_past, " min", "\r")
    }
    if(cnt_info == 1000000){
        cnt_info <- 0
        time_past <- round(difftime(Sys.time(), start_fit, 
                                    units = "mins"), 2)
        message("Tries: ", cnt_tries/1000000, " M", 
            " | Sharpe ratio: ", sharpe_ratio_fit, 
            " | Time past: ", time_past, " min", "\r")
        
    }
    cnt_total <- cnt_total + 1
}
finish_fit <- Sys.time()
spent <- difftime(finish_fit, start_fit, units = "mins")

# Storing the values in the table
fit_portfolio_values <- tibble(Return = port_fit_returns,
                           Risk = port_fit_risk,
                           SharpeRatio = sharpe_ratio_fit)


# Converting matrix to a tibble and changing column names
wts_fit <- as_tibble(rbind(wts_fit))
colnames(wts_fit) <- colnames(log_ret_xts)

# Combing all the values together
fit_portfolio_values <- tk_tbl(cbind(wts_fit, fit_portfolio_values), preserve_index = F)
```


Total fit trials : `r round(cnt_total/1000000)`M

```{r "simulation", echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# Simulation
start <- Sys.time()
for (i in seq_along(port_returns)) {
    
    wts <- runif(length(tick))
    wts <- round(wts/sum(wts),2)
    
    # Storing weight in the matrix
    all_wts[i,] <- wts
    
    # Portfolio returns
    
    port_ret <- sum(wts * mean_ret)
    port_ret <- ((port_ret + 1)^52) - 1
    
    # Storing Portfolio Returns values
    port_returns[i] <- round(port_ret,3)
    
    
    # Creating and storing portfolio risk
    port_sd <- sqrt(t(wts) %*% (cov_mat  %*% wts))
    port_risk[i] <- round(port_sd,3)
    
    # Creating and storing Portfolio Sharpe Ratios
    # Assuming 0% Risk free rate
    
    sr <- (port_ret-risk_free)/port_sd
    sharpe_ratio[i] <- round(sr,3)
}
finish <- Sys.time()
spent <- difftime(finish, start, units = "mins")

# Storing the values in the table
portfolio_values <- tibble(Return = port_returns,
                           Risk = port_risk,
                           SharpeRatio = sharpe_ratio)


# Converting matrix to a tibble and changing column names
all_wts <- tk_tbl(all_wts)

colnames(all_wts) <- colnames(log_ret_xts)

# Combing all the values together
portfolio_values <- tk_tbl(cbind(all_wts, portfolio_values))


portfolio_values <- portfolio_values %>%
    distinct(Risk, Return, .keep_all = TRUE)
```

Time spent on simulation: `r round(spent)` minutes

Portfolios amount: `r as.character(num_port/1000000)`M

Date range from `r min(index(log_ret_xts_compl))` to `r max(index(log_ret_xts_compl))`

```{r "single asset", echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# Simulation

wts_singles <- matrix(rep(0, length(tick)^2), 
                      nrow = length(tick), ncol = length(tick))
diag(wts_singles) <- 1

port_singles_returns <- vector('numeric', length = length(tick))

port_singles_risk <- vector('numeric', length = length(tick))

sharpe_ratio_singles <- vector('numeric', length = length(tick))

for (i in 1:length(tick)) {
    # Portfolio returns
    
    port_ret <- sum(wts_singles[i,] * mean_ret)
    port_ret <- ((port_ret + 1)^52) - 1
    
    # Storing Portfolio Returns values
    port_singles_returns[i] <- round(port_ret,3)
    
    
    # Creating and storing portfolio risk
    port_singles_sd <- sqrt(t(wts_singles[i,]) %*% (cov_mat  %*% wts_singles[i,]))
    port_singles_risk[i] <- round(port_singles_sd,3)
    
    # Creating and storing Portfolio Sharpe Ratios
    # Assuming 0% Risk free rate
    
    sr <- (port_ret-risk_free)/port_singles_sd
    sharpe_ratio_singles[i] <- round(sr,3)
}

# Storing the values in the table
single_portfolio_values <- tibble(Return = port_singles_returns,
                           Risk = port_singles_risk,
                           SharpeRatio = sharpe_ratio_singles)


# Converting matrix to a tibble and changing column names
wts_singles <- tk_tbl(wts_singles)

colnames(wts_singles) <- colnames(log_ret_xts)

# Combing all the values together
single_portfolio_values <- tk_tbl(cbind(wts_singles, single_portfolio_values))

single_portfolio_values <- single_portfolio_values %>%
    pivot_longer(1:(ncol(single_portfolio_values)-3),
             names_to = "Asset") %>%
    filter(value == 1) %>%
    select(-value)
    
```


## My current portfolio

```{r "my portfolio", echo = FALSE, warning = FALSE}
# Calculate the random weights

#wts_one <- c(0, 0.2751, 0.1958, 0.5290, rep(0,7))
#wts_one <- c(0.4583, 0.3162, 0.2255, rep(0,8))

wts_one <- rep(0, length(tick))
names(wts_one) <- tick
wts_one <- replace(wts_one, names(wts_have), wts_have)

# Calculate the portfolio returns
port_return_one <- (sum(wts_one * mean_ret) + 1)^52 - 1

# Calculate the portfolio risk
port_risk_one <- sqrt(t(wts_one) %*% (cov_mat %*% wts_one))

# Calculate the Sharpe Ratio
sharpe_ratio_one <- port_return_one/port_risk_one

portfolio_values_one <- tibble(Return = port_return_one,
                               Risk = port_risk_one,
                               SharpeRatio = sharpe_ratio_one)

# Converting matrix to a tibble and changing column names
wts_one <- as_tibble(rbind(wts_one))
colnames(wts_one) <- colnames(log_ret_xts)

# Combing all the values together
portfolio_values_one <- tk_tbl(cbind(wts_one, portfolio_values_one), preserve_index = F)

kbl(round(portfolio_values_one,2)) %>%
    kable_styling()
```


## Minimum Variance Portfolio Weights

```{r "minimum variance portfolio", echo = FALSE, out.width='100%'}
# The minimum variance portfolio
min_var <- portfolio_values[which.min(portfolio_values$Risk),]
p <- min_var %>%
    gather(names(min_var)[1]:names(min_var)[length(names(min_var))-3], 
           key = Asset,
           value = Weights) %>%
    mutate(Asset = as.factor(Asset)) %>%
    ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
    geom_bar(stat = 'identity') +
    theme_minimal() +
    labs(x = 'Assets', y = 'Weights', title = "Minimum Variance Portfolio Weights") +
    scale_y_continuous(labels = scales::percent) 

min_var <- min_var[,order(min_var[nrow(min_var),])]
min_var <- relocate(min_var, Return, Risk, SharpeRatio,.after = last_col())

kbl(round(min_var,2)) %>%
    kable_styling()


ggplotly(p)
```

## Tangency Portfolio Weights

```{r "tangency portfolio", echo = FALSE, out.width='100%'}
max_sr <- portfolio_values[which.max(portfolio_values$SharpeRatio),]

p <- max_sr %>%
    gather(names(max_sr)[1]:names(max_sr)[length(names(max_sr))-3], 
           key = Asset,
           value = Weights) %>%
    mutate(Asset = as.factor(Asset)) %>%
    ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
    geom_bar(stat = 'identity') +
    theme_minimal() +
    labs(x = 'Assets', y = 'Weights', 
         title = "Tangency Portfolio Weights") +
    scale_y_continuous(labels = scales::percent)

max_sr <- max_sr[,order(max_sr[nrow(max_sr),])]
max_sr <- relocate(max_sr, Return, Risk, SharpeRatio,.after = last_col())

kbl(round(max_sr,2)) %>%
    kable_styling()
ggplotly(p)
```

## Portfolio Sharpie ratio fit

```{r "portfolio fit res", echo=FALSE, out.width='100%'}
p <- fit_portfolio_values %>%
    gather(names(fit_portfolio_values)[1]:names(fit_portfolio_values)[length(names(fit_portfolio_values))-3], 
           key = Asset,
           value = Weights) %>%
    mutate(Asset = as.factor(Asset)) %>%
    ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
    geom_bar(stat = 'identity') +
    theme_minimal() +
    labs(x = 'Assets', y = 'Weights', 
         title = "Fit Portfolio Weights") +
    scale_y_continuous(labels = scales::percent)

kbl(round(fit_portfolio_values,2)) %>%
    kable_styling()

ggplotly(p)
```

```{r "efficient frontier", echo = FALSE, out.width='100%'}
# Visualizing all the portfolios

p <- portfolio_values %>%
    ggplot(aes(x = Risk, y = Return, color = SharpeRatio)) +
    geom_point() +
    theme_classic() +
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(labels = scales::percent) +
    #coord_cartesian(xlim = c(min(portfolio_values$Risk) - 0.02,
    #                         max(portfolio_values$Risk) + 0.02),
    #                ylim = c(min(portfolio_values$Return - 0.02),
    #                         max(portfolio_values$Return) + 0.02)) +
    labs(x = 'Annualized Risk',
         y = 'Annualized Returns',
         title = "Portfolio Optimization & Efficient Frontier") +
    geom_point(aes(x = Risk,
                   y = Return), data = min_var, color = 'red', size = 2) +
    geom_point(aes(x = Risk,
                   y = Return), data = max_sr, color = 'red', size = 2) +
    geom_point(aes(x = Risk, y = Return), 
               data = portfolio_values_one, color = "orange", size = 2) +
    geom_text(aes(x = Risk, y = Return, label = Asset),
              data = single_portfolio_values, 
              color = "darkgray", size = 3) +
    geom_point(aes(x = Risk, y = Return), data = fit_portfolio_values,
               color = "pink", size = 2)

p
#ggplotly(p)
```

Total time: `r round(difftime(Sys.time(), report_start_time, units = "mins"))` minutes

